Python as a machine learning language: experiences from the scikit-learn project

### 2013-06-16 paper.tex
section on extending scikit-learn
Emphasis is on avoiding frameworky code.
Feel free to excise the witty quote if you think it's not appropriate;
I just had to share it.

To ease code reuse, object-oriented techniques and design patterns are also used
all throughout the codebase. In particular, the \textit{duck typing} design
pattern is widely exploited to simplify implementation and skip the introduction
of superfluous classes. This technique indeed allows both for extensibility and
flexibility: user-defined estimators that follow our API and conventions should
be reusable (e.g., in a pipeline) as any built-in estimators, without them
actually inheriting from our base classes.

@@

>>>>

@@

\subsection{Extending \sklearn}

Any object that conforms to a small superset of the API
lined out in Section~\ref{sec:core-api},
(in particular, that implements methods \texttt{get\_params}
and \texttt{set\_params} for setting hyperparameters;
refer to the documentation at \url{http://scikit-learn.org} for details)
can be used as a \sklearn estimator
and be ``plugged into'' a pipeline, grid search or other meta-estimator
To define such objects, a user can either derive a new class
from \texttt{sklearn.base.BaseEstimator},
or define the entire class from scratch,
without even an \texttt{import sklearn}
in the relevant program module.
The Python principle of ``duck typing''
is employed to make sure that even instances of a class
that does not derive from any \sklearn class can be a first-class estimator.

In other places of the library,
in particular in the feature extraction and vectorization code
for text documents,
the toolkit is designed to be extensible.
Here, the estimators provide ``hooks'' for user-defined code:
objects or functions that provide a specific API
can be given as arguments at vectorizer construction time.
\sklearn code then calls into this code
to obtain the desired behavior.
Again, such user code can be kept agnostic
of the \sklearn class hierarchy,
operating on standard Python objects such as strings instead.

The design principle here is that user code
should not be tied to \sklearn,
which is a \textit{library}, not a \textit{framework}.
This principle avoids a variant of the well-known problem
with object-oriented design,
which is that users wanting a ``banana'' get
``a gorilla holding the banana and the entire jungle''
(Joe~Armstrong, interviewed by \citealp{seibel2009coders}:213):
a program that is not reusable with another machine learning toolkit
(or other solution to the problem at hand),
because it is intimately tied to \sklearn.
